---
title: "Publishable Graphs in R"
author: "Juan Ospina"
date: '2024-01-18'
output: 
  html_document: 
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

# First Part

## Load packages

```{r}
# Run the line below to install any missing packages
# install.packages("tidyverse")

# Packages
library(tidyverse) # to do almost everything in R. It's a package of packages. Includes dplyr, ggplot2, and many others
library(magrittr) # to pipe and assign
library(knitr) # to knit
library(psych) # to get descriptive statistics
library(psychTools) # to get the dataset that we'll use in the workshop
library(emmeans) # to get estimated means from models
library(jtools) # for summ(), replaces summary() for model stats
library(interactions) # for interaction plots
library(kableExtra) # to create HTML tables
```

## Get session info

This is useful when you need to run old code and figure out which version of the packages you were using

```{r}
sessionInfo()
```

## Custom functions

These functions simplify the code below

```{r}
# Format HTML Tables with Kable
kable_format <- . %>% 
  kable(format = "html") %>% 
  kable_styling(bootstrap_options = c("hover", "responsive"), font_size = 12, full_width = F, html_font = "helvetica")

# Get Percentages
factor_pct <- function(data, factor){
  data %>%
    count(!!!factor) %>%
    mutate(total = sum(n, na.rm = F),
           pct = round(n/total*100, 1)) %>%
    kable_format
}

# Get nice HTML table of descriptive statistics
describe_format <- function(data){
  data %>%
    describe %>%
    as.data.frame() %>%
    rownames_to_column(var = "var_name") %>%
    mutate_at(vars(-var_name), funs(round(., digits = 2))) %>%
    kable_format
}

# Descriptive Statistics by Factors
describe_factors_format <- function(data, factors, variables){
  data %>% 
    select(!!!variables) %>% 
    describeBy(factors, mat = TRUE, digits = 2) %>%
    kable_format
}
```


## Load dataset

Info about the dataset "affect" can be found here:

https://rdrr.io/cran/psychTools/man/affect.html
https://rdrr.io/cran/psychTools/man/msq.html

```{r}
# Get dataset. Note: need to install and load psychTools package
ds <- force(affect)
```

## Get variables names

```{r}
ds %>% 
  names() %>% 
  noquote()
```

## Clean variables

```{r}
ds %<>% # %<>% = ds <- ds %>%. With " %<>%", we're both piping (%>%) and assigning (<-) at the same time
  mutate(
    Study_num = ifelse(Study == "maps", 0, 1),
    
    # Compute new factor variable from the numerical one
    Study_f = factor(Study_num,
                    levels = c(0, 1),
                    labels = c("maps", "flat")),
    Film_f = factor(Film,
                    levels = c(1, 2, 3, 4),
                    labels = c("Frontline", "Halloween", "Nat Geo", "Parenthood")),
    Film_New = ifelse(Film == 1, 0,
                      ifelse(Film == 2, 1, NA)),
    Film_New_f = factor(Film_New,
                    levels = c(0, 1),
                    labels = c("Frontline", "Halloween")),
    Film_Front = ifelse(Film == 1, 0, 1),
    Film_Front_f = factor(Film_Front,
                    levels = c(0, 1),
                    labels = c("Frontline", "Other Movies")),
    
    # Rename variables to more intuitive names while keeping the original variables
    extraver = ext,
    neurot = neur,
    EnergeticArousal2 = EA2,
    TenseArousal2 = TA2,
    PositiveAffect2 = PA2,
    NegativeAffect2 = NA2,
    
    # Standardize variables
    Zextraver = scale(extraver, center = TRUE, scale = TRUE),
    Zextraver = as.numeric(Zextraver),
    Zneurot = scale(neurot, center = TRUE, scale = TRUE),
    Zneurot = as.numeric(Zneurot),
    ZEnergeticArousal2 = scale(EnergeticArousal2, center = TRUE, scale = TRUE),
    ZEnergeticArousal2 = as.numeric(ZEnergeticArousal2),
    ZTenseArousal2 = scale(TenseArousal2, center = TRUE, scale = TRUE),
    ZTenseArousal2 = as.numeric(ZTenseArousal2),
    ZPositiveAffect2 = scale(PositiveAffect2, center = TRUE, scale = TRUE),
    ZPositiveAffect2 = as.numeric(ZPositiveAffect2),
    ZNegativeAffect2 = scale(NegativeAffect2, center = TRUE, scale = TRUE),
    ZNegativeAffect2 = as.numeric(ZNegativeAffect2),
  )

# To double-check that factors were coded correctly
factor_pct(ds, quos(Film, Film_f, Film_New, Film_New_f, Film_Front, Film_Front_f))
factor_pct(ds, quos(Study, Study_num, Study_f))

# Create list of continuous variables
vars_cont <-
  quos(
    extraver, 
    neurot, 
    EnergeticArousal2,
    TenseArousal2,
    PositiveAffect2,
    NegativeAffect2,
    Zextraver, 
    Zneurot, 
    ZEnergeticArousal2,
    ZTenseArousal2,
    ZPositiveAffect2,
    ZNegativeAffect2
  )

# Get descriptive statistics of continuous variables
ds %>%
  select(
    !!!vars_cont) %>%
  describe_format()

# Get descriptive statistics of continuous variables by factors
describe_factors_format(ds, list(ds$Film_f), vars_cont)
describe_factors_format(ds, list(ds$Film_f, ds$Study_f), vars_cont)

```

# Second Part

## Settings plots

```{r}
# remove anoying labels from the graph
guides_no <- guides(color = FALSE, fill = FALSE)

# black and white theme
theme_overall <- theme_bw()

# many more changes that you can make to the graphs
theme_specifics <- 
  theme(
    strip.text = element_text(color = "White"), 
    strip.background = element_rect(fill = "black", colour = "black", size = NULL, linetype = NULL, inherit.blank = FALSE), 
    axis.text = # change y-axis font
      element_text(colour = "black", size = 16, family = "Times New Roman"),
    axis.text.x = # change size of x-axis font
      element_text(colour = "black", size = 14, family = "Times New Roman"),
    legend.title = # change legend title font
      element_text(colour = "black", size = 16, family = "Times New Roman"),
    legend.text = # change legend labels font
      element_text(colour = "black", size = 16, family = "Times New Roman"),
    axis.title = element_text(colour = "black", size = 22, family = "Times New Roman"), 
    strip.text.x = element_text(size = 22, colour = "white", family = "Times New Roman"),
    plot.title = element_text(size = 22, hjust = 0.5, family = "Times New Roman"),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    plot.caption = element_text(size = 14, hjust = 0),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # axis.title.x = element_blank(),
    axis.ticks = element_line(colour = "black", size = 0.8),
    # panel.border = element_rect(colour = "black"),
    # axis.ticks.x = element_blank(),
    # legend.position = 'none'
    # legend.key = element_rect(size = 10),
    # legend.key.size = unit(1, 'lines')
  )

# Colors to choose from: https://sape.inf.usi.ch/quick-reference/ggplot2/colour

# Colors and labels of IVs
scale_colour_movies <- 
  scale_colour_manual(
    values = c("Frontline" = "palegreen4", "Halloween" = "steelblue3", 
               "Nat Geo" = "salmon", "Parenthood" = "gold3"),
    name = "Leader",
    breaks=c("Frontline", "Halloween", "Nat Geo", "Parenthood"),
    labels = c("Frontline" = "Frontline", "Halloween" = "Halloween", 
               "Nat Geo" = "National\nGeographic", "Parenthood" = "Parenthood"))

scale_x_movies <- 
  scale_x_discrete(breaks = c("Frontline", "Halloween", "Nat Geo", "Parenthood"), 
                   labels = c("Frontline", "Halloween", "National\nGeographic", "Parenthood"))

colors_studies <- c("bisque4", "orchid3")

# Means text for bar graphs
geom_text_bar <- geom_text(aes(label = round(emmean, digits = 2), y = 1.2), size = 7, color = "black", family = "Times New Roman")

# Settings for standardized graphs
breaks_std <- c(-1, -0.5, 0, 0.5, 1)
labels_std <- c("-1SD", "-0.5", "0", "0.5", "+1SD")
```

## Bar Plots

```{r}
# Estimate the model with fixed effects and store results in Model1
Model1 <- lm(PositiveAffect2 ~  # Outcome
                 Film_f + extraver, #IV
             data = ds)

# summary model with confidence intervals
summ(Model1, confint = TRUE, ci.width = .95, digits = 3)

# To get estimated means from model
means <- emmeans(Model1, specs = "Film_f")
means

# Get all contrasts
pairs(means, reverse = TRUE, adjust = "none")

# Transform the "means" object to a dataframe
means_graph <- as.data.frame(means)

# If you would like to add the labels of the scale, you can use the code below and in the graph
# ScaleLabels <- c("1\nNever", "2\nRarely", "3\nSometimes", "4\nOften", "5\nAlways")

# Get publishable bar graphs from the model
means_graph %>% 
  ggplot(aes(x = Film_f, y = emmean, color = Film_f, group = Film_f)) + 
  geom_segment(aes(x = Film_f, xend = Film_f, y = -10, yend = emmean), size = 35) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), 
                width = .3, position = position_dodge(0.1), color = "black") +
  geom_text_bar +  
  coord_cartesian(ylim = c(0, 29)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25, 29)) +
                     # labels = ScaleLabels) +
  labs(title = "Post Positive Affect by Movie",
       y = "Post Positive Affect",
       x = "Movie") +
  theme_overall + 
  theme_specifics +
  guides_no +
  scale_colour_movies + 
  scale_x_movies

# To save plots
ggsave(file = "figures/Movies on Post Positive Affect.jpg", width = 7, height = 5, dpi = 1000)
```

## Moderation by Categorical Predictors

```{r}
# Estimate the model with fixed effects and store results in Model1
Model1 <- lm(PositiveAffect2 ~  # Outcome
                 Film_f * Study_f + extraver, # IVs
             data = ds)

summ(Model1, confint = TRUE, ci.width = .95, digits = 3, part.corr = TRUE, vifs = TRUE)

means <- emmeans(Model1, specs = "Film_f", by = "Study_f")
means

# Get contrasts by IV2
pairs(means, reverse = TRUE, adjust = "none")

means <- emmeans(Model1, specs = "Study_f", by = "Film_f")
means

# Get contrasts by IV1
pairs(means, reverse = TRUE, adjust = "none")

# ScaleLabels <- c("1\nLowest rung", "2", "3", "4", "5", "6", "7", "8", "9", "10\nHighest rung")

cat_plot(Model1, 
         pred = Film_f, 
         pred.labels = c("Frontline", "Halloween", "National\nGeographic", "Parenthood"),
         modx = Study_f,
         legend.main = "Study",
         modx.labels = c("MAPS", "FLAT"),
         colors = colors_studies,
         geom = "bar",
         interval = TRUE,
         int.type = "confidence",
         int.width = .95) +
  theme_overall +
  theme_specifics +
  coord_cartesian(ylim = c(0, 29)) +
  # scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), labels = ScaleLabels) +
  labs(title = "Post Positive Affect\nby Movie and Study", 
       y = "Post Positive Affect",
       x = "Movie")

# To save plots
ggsave(file = "figures/Movies and Study on Post Positive Affect.jpg", width = 7, height = 5, dpi = 1000)
```

## Moderation by Continuous Predictors

```{r}
# Estimate the model with fixed effects and store results in Model1
Model1 <- lm(ZPositiveAffect2 ~  # Outcome
                 Zextraver * Zneurot, # Continuous IVs
             data = ds)

summ(Model1, confint = TRUE, ci.width = .95, digits = 3, part.corr = TRUE, vifs = TRUE)

# Simple slopes code
# https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html
sim_slopes(Model1, pred = Zextraver, modx = Zneurot, johnson_neyman = FALSE, digits = 3)

sim_slopes(Model1, pred = Zneurot, modx = Zextraver, johnson_neyman = FALSE, digits = 3)

# Colors of slopes
colors_neurot <- c("gold3", "salmon")

interact_plot(
  Model1, 
  pred = Zneurot, 
  modx = Zextraver,
  colors = colors_neurot,
  modx.values = "plus-minus",
  # plot.points = TRUE,
  interval = TRUE,
  int.type = "confidence",
  int.width = .95,
  x.label = "Neuroticism",
  y.label = "Post Positive Affect",
  main.title = "Post Positive Affect\nby Neuroticism and Extraversion",
  legend.main = "Extraversion") +
  theme_overall +
  theme_specifics +
  theme(legend.position = c(.85, .85)) +
  coord_cartesian(ylim = c(-1, 1), xlim = c(-1, 1)) +
  scale_y_continuous(breaks = breaks_std, labels = labels_std) +
  scale_x_continuous(breaks = breaks_std, labels = labels_std)

# To save plots
ggsave(file = "figures/Neuroticism and Extraversion on Post Positive Affect.jpg", width = 7, height = 5, dpi = 1000)
```

```{r}
# Estimate the model with fixed effects and store results in Model1
Model1 <- lm(ZPositiveAffect2 ~  # Outcome
                 Zextraver * Zneurot + ZNegativeAffect2, # Continuous IVs
             data = ds)

summ(Model1, confint = TRUE, ci.width = .95, digits = 3, part.corr = TRUE, vifs = TRUE)

# Simple slopes code
# https://cran.r-project.org/web/packages/interactions/vignettes/interactions.html
sim_slopes(Model1, pred = Zextraver, modx = Zneurot, johnson_neyman = FALSE, digits = 3)

sim_slopes(Model1, pred = Zneurot, modx = Zextraver, johnson_neyman = FALSE, digits = 3)

# Colors of slopes
colors_neurot <- c("gold3", "salmon")

interact_plot(
  Model1, 
  pred = Zneurot, 
  modx = Zextraver,
  colors = colors_neurot,
  modx.values = "plus-minus",
  # plot.points = TRUE,
  interval = TRUE,
  int.type = "confidence",
  int.width = .95,
  x.label = "Neuroticism",
  y.label = "Post Positive Affect",
  main.title = "Post Positive Affect\nby Neuroticism and Extraversion",
  legend.main = "Extraversion") +
  theme_overall +
  theme_specifics +
  theme(legend.position = c(.85, .85)) +
  coord_cartesian(ylim = c(-1, 1), xlim = c(-1, 1)) +
  scale_y_continuous(breaks = breaks_std, labels = labels_std) +
  scale_x_continuous(breaks = breaks_std, labels = labels_std)

# To save plots
ggsave(file = "figures/Neuroticism and Extraversion on Post Positive Affect.jpg", width = 7, height = 5, dpi = 1000)
```
